Wired for Success: Sweden's Electrical Innovators in the Age of Industrial Transformation

### 2023-05-14 index.qmd
add slideshow

---
title: "Wired for Success: Sweden's Electrical Innovators in the Age of Industrial Transformation"
title-slide-attributes:
  data-background-image: "assets/preview.png"
  data-background-size: "30%"
  data-background-opacity: "0.6"
format: 
  revealjs:
    code-overflow: wrap
    width: 1600
    height: 900
    fig-align: center
    default-image-extension: svg
    footer: "[{{< fa house >}}](https://j-jayes.github.io/Data-Visualization-in-R-Lund/)"
image: "preview_img/03-coordinate-systems-and-axes.jfif"
favicon: html/favicon.ico
---

@@

::: {.columns}

@@

We get biographies and career trajectories of 69448 individuals!

8 of the 10 volumes are digitized by librarians in Uppsala – thank you <3

@@

::: {.column width="10%"}

@@

::: {.column width="20%"}

@@

>>>>

---
title: "Wired for Success: Sweden's Electrical Innovators in the Age of Industrial Transformation"
title-slide-attributes:
  data-background-image: "assets/preview.png"
  data-background-size: "30%"
  data-background-opacity: "0.6"
format: 
  revealjs:
    code-overflow: wrap
    width: 1600
    height: 900
    fig-align: center
    default-image-extension: svg
    footer: "[{{< fa house >}}](https://j-jayes.github.io/Data-Visualization-in-R-Lund/)"
    slide-number: c
image: "preview_img/03-coordinate-systems-and-axes.jfif"
favicon: html/favicon.ico
---

@@

@@

- Biographies and career trajectories of ~ 69,448 individuals!

- 8 of the 10 volumes are digitized by librarians in Uppsala – thank you <3

@@

::: {.column width="5%"}

@@

::: {.column width="25%"}

@@

![](assets/tidy.jpg)

🧑
📝
🗑️

Arrow

💻🧹

Arrow

📚
🎓
💼

## NLP challenges

::: {.column width="50%"}

<br>

- Many abbreviations

- Similar structure for each entry but not exactly the same information in the same order

::: {.column width="50%"}

<br>

![](assets/Karl_Lund_crop.png){fig-align="center"}

## 1. Translate

```{css}
code.sourceCode {
  font-size: 1.3em;
  /* or try font-size: xx-large; */
}
```

```{python}
#| eval: false
#| echo: true

def translate_and_structure_text(swedish_text):
    try:
        # Translate the Swedish text to English
        translate_prompt = f"Translate the following abbreviated Swedish biography to English: {swedish_text}"
        translation_response = openai.ChatCompletion.create(
            model="gpt-3.5-turbo",
            messages=[
                {"role": "system", "content": "You are an expert on Swedish family history."},
                {"role": "user", "content": f"{translate_prompt}"}
            ]
        )

## 2. Structure

```{python, eval=FALSE, echo=TRUE}

structure_prompt = f"Given the original Swedish biography: {swedish_text}\nAnd its English translation: {english_text}, Structure the biography in Schema.org/Person format as a JSON object. Include dates wherever possible. Only provide a  RFC8259 compliant JSON response."

structure_response = openai.ChatCompletion.create(
    model="gpt-3.5-turbo",
    messages=[
        {"role": "system", "content": "You are an expert on Swedish family history and the Schema.org/Person format."},
        {"role": "user", "content": f"{structure_prompt}"}
    ]
)
```

## 3. Augment

```{python, eval=FALSE, echo=TRUE}

def geocode_birthplace(birthplace, geolocator, geocoded_cache):
    if isinstance(birthplace, dict):
        if "name" in birthplace:
            birthplace_name = birthplace["name"]
        elif "address" in birthplace and "addressLocality" in birthplace["address"]:
            birthplace_name = birthplace["address"]["addressLocality"]
        else:
            print(f"Invalid birthplace data: {birthplace}")
            return None
    elif isinstance(birthplace, str):
        birthplace_name = birthplace
    else:
        print(f"Invalid birthplace data: {birthplace}")
        return None

    if birthplace_name not in geocoded_cache:
        try:
            location = geolocator.geocode(birthplace_name)
            if location:
                geocoded_cache[birthplace_name] = {
                    "latitude": location.latitude,
                    "longitude": location.longitude
                }
            time.sleep(1)  # Add a 1-second delay between geocoding requests
        except Exception as e:
            print(f"Error geocoding birthplace {birthplace_name}: {e}")

    return geocoded_cache.get(birthplace_name)
```

## Output

```{python, eval=FALSE, echo=TRUE}
{
    "original": "Axell, Folke Adolf, civilingenjör, Norrköping, f. 4/12/08 i Ljusnarsberg, Örebro 1. G. 36 m. Marianne Hammarberg. Barn: Anita f. 37, Peter 42. — Avg:ex. fr. tekn. gymn. i Örebro 28, civ:ing:ex. v. KTH 34. Ing:assist. v. telegr :v. i Nässjö 35, linjeing. v. te-legr :v. i Norrköp. sed. 46. Led. av Sv. tekn :fören. sed. 35.",
    "translated": "Axell, Folke Adolf, engineer, from Norrköping, born on December 4th, 1908 in Ljusnarsberg, Örebro 1. Married in 1936 to Marianne Hammarberg. Children: Anita born in 1937, Peter born in 1942. He graduated from the Technical Gymnasium in Örebro in 1928, received his engineering degree from KTH in 1934. Worked as an engineering assistant at the telegraph station in Nässjö in 1935, and as a line engineer at the telegraph station in Norrköping since 1946. Member of the Swedish Technical Association since 1935.",
    "structured": {
        "@context": "https://schema.org",
        "@type": "Person",
        "name": "Folke Adolf Axell",
        "birthDate": "1908-12-04",
        "birthPlace": {
            "name": "Ljusnarsberg, Örebro",
            "latitude": 59.916667,
            "longitude": 15.0
        },
        "jobTitle": "Engineer",
        "memberOf": {
            "@type": "Organization",
            "name": "Swedish Technical Association"
        },
        "children": [
            {
                "@type": "Person",
                "name": "Anita",
                "birthDate": "1937"
            },
            {
                "@type": "Person",
                "name": "Peter",
                "birthDate": "1942"
            }
        ],
        "spouse": {
            "@type": "Person",
            "name": "Marianne Hammarberg",
            "marriageDate": "1936"
        },
        "education": [
            {
                "@type": "EducationalOccupationalCredential",
                "credentialCategory": "Vocational",
                "issuer": {
                    "@type": "Organization",
                    "name": "Technical Gymnasium in Örebro"
                },
                "endDate": "1928"
            },
            {
                "@type": "EducationalOccupationalCredential",
                "credentialCategory": "Degree",
                "issuer": {
                    "@type": "Organization",
                    "name": "KTH"
                },
                "endDate": "1934"
            }
        ],
        "worksFor": {
            "@type": "Organization",
            "name": "Telegraph station in Norrköping",
            "startDate": "1946"
        },
        "hasOccupation": [
            {
                "@type": "Occupation",
                "name": "Engineering assistant",
                "employer": {
                    "@type": "Organization",
                    "name": "Telegraph station in Nässjö"
                },
                "startDate": "1935",
                "endDate": "1935"
            },
            {
                "@type": "Occupation",
                "name": "Line engineer",
                "employer": {
                    "@type": "Organization",
                    "name": "Telegraph station in Norrköping"
                },
                "startDate": "1946"
            }
        ]
    }
}
```
